---
title: "2019 Nature PhD Students Survey"
authors: "Icíar Fernández, Jacob Gerlofs"
date: "26/02/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: yes
  editor_options:
    chunk_output_type: console 
---

## Introduction

For the past five years, the iconic science journal Nature has launched a survey for PhD students in STEM fields to share their experience in graduate school, hoping to illuminate the goals, challenges, and sources of satisfaction for doctoral students across seven continents. Last year's survey collected data from over 6000 graduate students, which constitutes the highest response rate in the survey's history. The [full data](https://figshare.com/s/74a5ea79d76ad66a8af8) from the survey was made publicly available following publication of an [article](https://www.nature.com/articles/d41586-019-03459-7) discussing the results. It is interesting to note that the survey was offered in English, Spanish, Chinese, French, and Portuguese - open-form questions have not been translated to English if answered by the participant in another language. Available materials include anonymysed raw data, the questionnaire that was provided to PhD students, and a presentation of the survey data. 

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(dplyr)
library(plyr)
library(kableExtra)
library(here)
```

## Data description

According to the script with survey information that was provided, there were a total of 65 questions. Not all questions were mandatory, and there was a mix of single choice (yes/no), multiple choice (several options) and free-form questions. 

In the dataset, each row represents an individual who participated in the survey, whereas each row represents a question. We have noticed some redundancy in the dataset column that will require substantial cleanup of the data as part of our next project milestone. For instance, Q12 ("What prompted you to study outside your country of upbringing?") was presented in the survey as a multiple choice question with 11 possible answers (a-k), with the last one (k) being open-form ("If other, please specify"). In the data frame, 11 rows correspond to Q12, each one composed of 2 values: NA, and 1/11 possible answers. As such, the column named Q12_1 only contains NA values and answer "(a) To study at a specific university"; whereas Q12_2 only contains NA values and answer "(b) Lack of funding opportunities in my home country", and so on. We plan on combining columns Q12_1:Q12_11 into a single Q12 column using dplyr::coalesce(), following the same rationale for other redundant columns in the dataset. In addition, open-form questions such as (k) in this specific example will be dropped due to the difficulty in analyzing this, and the fact that they contain answers in different languages.

```{r survey-dataset, warning=FALSE}
survey_data <- readxl::read_xlsx(here::here("docs", "Nature_PhD_Survey.xlsx"))
dim(survey_data)
```

Due to this redundancy, the dimensions of the raw dataset when downloaded are 6812 rows (participants) by 274 columns (questions), whereas the actual survey only has 63 questions. Below is the complete list of questions, which is a simplified version of the Word document provided [here](https://figshare.com/s/74a5ea79d76ad66a8af8), which includes all the possible answers for each question. For simplicity, we have only included the question, its type, and the category it belongs to within the survey.

```{r survey-questions, echo=FALSE, warning=FALSE, message=FALSE}
survey_questions <- readxl::read_xlsx(here::here("docs", "survey_questions.xlsx"))
dim(survey_questions) # 63 (questions) by 4 columns (question number, category, question, and type)
knitr::kable(survey_questions) %>% kable_styling(bootstrap_options = "condensed", full_width=F, fixed_thead=TRUE) # print
```

## Exploring the dataset
```{r basic demographics, echo=FALSE, warning=FALSE, message=FALSE}

survey_demographics <- survey_data[-1, ] #removing first row of data file, which contains question label and not participant data.

ggplot(survey_demographics, aes(Q5)) +
  geom_bar(width = .8, fill = "steelblue") +
  coord_flip() +
  xlab("Continent") +
  ylab("Where do you currently live?") +
  theme_minimal()
```
It would appear that most of the participants in this survey are from Europe, the Americas, or Asia. This is not very surprising, but is good to verify the sample is relatively representative of the distribution of higher education institutes around the world.

## Plan of Action




## References 


